---
title: "Core Value Drivers Reporting"
format: html
---

# Internal Fill Rate

- Pull “RRX – Employee Demographics” Report in Workday
  - Effective as of Date: Last date in period of interest (e.g., 1/31/2025)
  - Organization: Regal Rexnord Team (Louis Pinkham)
- Pull "RRX - 9Box" Report in Workday *NEED ACCESS*
  - Supervisory Organization: Regal Rexnord Team (Louis Pinkham)
  - Include Subordinate Organizations
  - Click hyperlink "Toal" above visual 9Box
  - Export to Excel

```{r}
#| warning: false
library(tidyverse)

# specify period
start_date <- as.Date("2025-01-01")
end_date <- Sys.Date()

# active employees
employment_demographics <- readxl::read_xlsx("inputs/RRX - Employment Demographics.xlsx", 
                                             skip = 12) |> 
  janitor::clean_names()

# temp time in position
time_in_position <- readxl::read_xlsx("inputs/temp-time in role.xlsx", skip = 17) |> 
  janitor::clean_names() |> 
  select(employee_id, time_in_job_profile_start_date)

# add time in position to main file
df <- employment_demographics |> 
  left_join(time_in_position, by = c("sso_number" = "employee_id"))

# all job changes to compare
job_changes <- readxl::read_xlsx("inputs/RRX_-_Worker_Data_-_Hires_and_Job_Changes.xlsx", skip = 7) |> 
  janitor::clean_names() |> 
  filter(str_detect(business_process_reason, regex("hire employee|promotion|lateral move", ignore_case = TRUE)))


```

- Add column [Internal/External]
  - If Hire Date = [Time in Job Profile Start Date] then "External"
  - If Hire Date < [Time in Job Profile Start Date] then "Internal"
  
```{r}
df <- df |> 
  mutate(internal_external = case_when(
    hire_date == time_in_job_profile_start_date ~ "External",
    hire_date < time_in_job_profile_start_date ~ "Internal",
    TRUE ~ "Error"
  ))
```

- Filter on [Time in Job Profile Start Date]
  - (1/1/{yyyy}) < [Time in Job Profile Start Date] < Last day in period
  - Feb update example: 1/1/2025 < [Time in Job Profile Start Date] < 1/31/2025
  
```{r}
df <- df |> 
  filter(time_in_job_profile_start_date > start_date & time_in_job_profile_start_date < end_date)
```

- Filter on [Career Level]
  - Career Level != M1
  
```{r}
df_manager <- df |> 
  filter(!career_level %in% c("AT1", "AT2", "AT3", "M1") & !is.na(career_level))
```

  
TO DO:
- Add Time in Job Profile Start Date to reports (Potentially WQL Alias: timeInJobProfileStartDate)
- Get access to RRX - 9Box Report (time in job profile)
- Clarify career_levels to exclude