---
title: "Internal Fill Rate"
format: html
editor: visual
---

### Load data sources

-   People Analytics - Active and Terminated Report (PA - AT Report)

-   RRX\_-\_Worker_Data\_-\_Hires_and_Job_Changes Report (Movement Report)

    
### Definitions

- Period:

    -   Start date: First date of current year

    -   End date: Last date of previous month
    
- internal_external:

    -   External: hire date is the same as job profile start date

    -   Internal: job profile start date is after hire rate

### Filter

-   Period: Job profile start date is within the period of interest

-   Manager+: remove career levels starting with "AT", "P", "M1", and blanks

### Add additional context from movement report

-   Remove duplicate rows for promotions (remove business process type promote employee outbound)

-   Matches when profile start date and effective date are the same

```{r}
library(tidyverse)

#| label: set_up

# specify period: first date of current year and last day of previous month
start_date <- "2024-01-01"
end_date <- "2024-12-31"

# define filters to get target population
# manager+
# employees only no contingent
apply_filters <- function(df) {
  df |> 
    filter(!career_level %in% c("AT1", "AT2", "AT3", "AT4", 
                              "P1", "P2", "P3", "M1") & 
           !is.na(career_level)) |> 
    filter(worker_type == "Employee")
}

# set the grouping variables of interest
grouping_vars <- list(NULL, 
                      "segment_function", 
                      c("segment_function", "division_function"), 
                      c("segment_function", "division_function", "location"),
                      c("segment_function", "career_level_bucket"))

# set name for output sheet name for each grouping variable
output_sheet_names <- c("Overall",
                        "Segment",
                        "Segment by Division",
                        "Division by Location",
                        "Segment by Level")

# check if the number of grouping variables matches the number of sheet names
if (length(grouping_vars) != length(output_sheet_names)) {
  stop("The number of grouping variables and sheet names must be the same.")
}
```



```{r}
#| label: pa_at

# file path for latest active and terminated
file_path_pa_at <- "inputs/People Analytics - AT 2024.xlsx"

# add career_level and internal/external flag
df_pa_at <- readxl::read_xlsx(file_path_pa_at, skip = 1) |> 
    janitor::clean_names() |>
  mutate(career_level_bucket = factor(case_when(
    career_level %in% c("E2", "E3", "E4", "E5") ~ "VP",
    career_level %in% c("M5", "E1") ~ "Director",
    career_level %in% c("M2", "M3", "M4") ~ "Manager",
    career_level == "M1" ~ "Supervisor",
    career_level %in% c("P1", "P2", "P3", "P4", "P5", "P6") ~ "Professional",
    career_level %in% c("AT1", "AT2", "AT3", "AT4") ~ "Administrative_Technical",
    is.na(career_level) ~ "None",
    TRUE ~ "ERROR_unmapped_level"),
    levels = c("VP", "Director", "Manager", "Supervisor", "Professional", "Administrative_Technical", "None", "ERROR_unmapped_level"),
    ordered = TRUE)) |>
  mutate(internal_external = case_when(
    hire_date == time_in_job_profile_start_date ~ "External",
    hire_date < time_in_job_profile_start_date ~ "Internal",
    TRUE ~ "Error"))

# filter data to current period
# add project filters to get target population
pa_at <- df_pa_at |> 
    filter(time_in_job_profile_start_date >= start_date &
           time_in_job_profile_start_date <= end_date)
```

```{r}
#| label: worker_changes

# file path to most recent changes report
file_path_worker_changes <- "inputs/RRX_-_Worker_Data_-_Hires_and_Job_Changes 2024.xlsx"

# read in worker changes
df_worker_changes_raw <- readxl::read_xlsx(file_path_worker_changes, skip = 7) |> 
  janitor::clean_names() 

df_worker_changes <- df_worker_changes_raw |> 
  select(employee_id, effective_date, business_process_type, business_process_reason, starts_with("prior_"), starts_with("new_")) 

worker_changes <- df_worker_changes |>
  filter(business_process_type != "Promote Employee Outbound") 
```



```{r}
#| label: join_data

# join data together matching on employee id and effective date/job profile start date
# apply filters to get target population
full_df <- pa_at |> 
  left_join(worker_changes, by = join_by("employee_id", "time_in_job_profile_start_date" == "effective_date")) |> 
  apply_filters()

# validation: check if many-to-many caused from multiple moves
qa_multiple_moves <- full_df %>%
  group_by(employee_id) %>%
  filter(n() > 1) %>%
  ungroup()
```



```{r}
#| label: data_to_share

# all data to be shared with results
full_df_scrubbed <- full_df |> 
  mutate(effective_date = case_when(
    is.na(business_process_reason) ~ NA,
    TRUE ~ time_in_job_profile_start_date)) |> 
  select(employee_id, worker, hire_date, time_in_job_profile_start_date, internal_external, 
         segment_function, division_function, location, career_level, 
         effective_date, business_process_type, business_process_reason, 
         prior_job_profile, prior_career_level, prior_comp_grade, 
         new_job_profile, new_career_level, new_comp_grade) |> 
  mutate_at(vars(contains("date")), as.Date) |> 
  arrange(time_in_job_profile_start_date)
```


```{r}
#| label: function_calculate_internal_fill

calculate_internal_fill <- function(df = full_df, grouping_var) {
  df <- df |> 
    group_by(across(all_of(grouping_var)), internal_external) |> 
    summarize(count = n(), .groups = "drop_last") |> 
    group_by(across(all_of(grouping_var))) |> 
    complete(internal_external = c("Internal", "External"), 
             fill = list(count = 0)) |>  
    mutate(total = sum(count),
           fill_rate = round(count / total, 2)) |> 
    ungroup() |> 
    filter(internal_external == "Internal") |> 
    arrange(desc(total)) |> 
    select(-internal_external) |> 
    rename('Roles Filled Internally' = count,
           'Total Filled Roles' = total,
           'Internal Fill Rate' = fill_rate)
  return(df)
}
```



```{r}
#| label: get_output

# run IFR function for each grouping variable
results <- map(grouping_vars, ~ calculate_internal_fill(full_df, .x))

# name each grouping var output to its sheet name
results_named <- set_names(results, output_sheet_names)

# add the scrubbed data details 
results_named[["Data Details"]] <- full_df_scrubbed

# name the output excel file
output_name <- paste0("outputs/Internal Fill Rate ", start_date, " to ", end_date, ".xlsx")
```



```{r}
# use openxlsx to format output as tables within excel
library(openxlsx)

wb <- createWorkbook()

for (sheet_name in names(results_named)) {
  addWorksheet(wb, sheet_name)
  writeDataTable(wb, sheet = sheet_name, x = results_named[[sheet_name]], tableStyle = "TableStyleMedium2")
}

saveWorkbook(wb, output_name, overwrite = TRUE)
```


### Validation Checks

```{r}
# on job changes, not diff profile/hire date
# excluding terminations and outbound to remove duplicates 
qa_change_dates_same <- df_worker_changes |> 
  filter(business_process_type != "Promote Employee Outbound") |> 
  filter(!new_career_level %in% c("AT1", "AT2", "AT3", "AT4", 
                                  "P1", "P2", "P3", "M1") & 
           !is.na(new_career_level)) |> 
  filter(!str_detect(business_process_reason, "Terminate")) |> 
  anti_join(pa_at, by = "employee_id")

qa_change_dates_same |> 
  group_by(business_process_reason) |> 
  summarize(count = n())
```
